---
title: "Getting started with ocular"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Getting started with ocular}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(ocular)
```

`ocular` retrieves multispectral satellite data and calculates spectral
indices for crop monitoring and agricultural research, including vegetative
indices (VI). This vignette follows one query from image retrieval to a
delineated raster and a time series of VI values.

Remote examples are not run when the vignette is built. They query the
Microsoft Planetary Computer Data Catalog and should be run where network
access is available.

## The core workflow

A query needs a point within the targeted crop production zone, a date range,
and the dimensions of the surrounding analysis area. Coordinates are supplied
in WGS 84. At least one of `x_metres` and `y_metres` is required; when one is
omitted, `ocular` uses the supplied dimension for both.

```{r retrieve-scenes, eval=FALSE}
rs <- get_rs(
  longitude = -96.80342,
  latitude = 33.26425,
  start_date = "2005-03-15",
  end_date = "2005-09-05",
  x_metres = 1000,
  index_name = "NDVI",
  source = "landsat-5"
)

rs
```

`get_rs()` returns an `ocular` object containing the query, the retrieved STAC
scenes, and the parameter settings that subsequent functions will use. The
supplied point is also the reference, or seed, for field boundary delineation.

Run the default delineation once, then reuse the result for each required
output.

```{r delineate-and-convert, eval=FALSE}
field <- boundary_delineation(rs)

vi_raster <- as_raster(field)
vi_series <- as_time_series(field)

vi_raster
head(vi_series)
```

`boundary_delineation()` stores the resulting field mask on the `ocular`
object. By default, `as_raster()` calculates the median value for each pixel
across the retrieved scenes. `as_time_series()` calculates the mean value
within the field for each day; `time_aggregate` can instead group results by
month or year.

`ocular` field boundary delineation minimises the need for labelled training
data by using the supplied point as a reference (seed) for the retrieved
multi-temporal spectral feature stack. It does not classify crop production
status, and its result should be checked for the intended location and use.

If delineation returns an empty mask, raster values are `NA` and the time
series has no rows. `ocular` does not replace an empty result with the full
analysis area.

## Retrieval choices

### Image source

The explicit source above keeps the example within Landsat 5/7. With
`source = "auto"`, the start date determines the collection: Sentinel-2 from
1 July 2015, Landsat 8/9 from 1 April 2013, and Landsat 5/7 for earlier dates.
Set the source explicitly when a study requires a consistent sensor family.

### Scene filtering and pixel masking

`max_cloud_cover` sets a scene-level cloud-cover ceiling using STAC metadata.
It does not mask individual pixels. The optional `scl_classes` argument
excludes sensor-specific Sentinel-2 SCL class values or Landsat `QA_PIXEL` bit
positions. Its default, `NULL`, applies no per-pixel cloud or quality mask.
Consult `?get_rs` for the accepted ranges and choose exclusions appropriate to
the source and analysis rather than transferring class values between sensors.

### Output and delineation indices

`index_name` controls the VI values returned by `as_raster()` and
`as_time_series()`. Field boundary delineation uses `search_index`, which is
`"EVI2"` by default. The two can be set independently. Supply non-default
delineation settings through `params` at retrieval so that they are stored on
the object.

`search_start_date` and `search_end_date` can restrict the scenes used to
construct boundary evidence. They must fall within the retrieval period and do
not remove dates from the output series.

## Working with the analysis area

Calling an output function before delineation returns values for the full
analysis area:

```{r full-analysis-area, eval=FALSE}
analysis_raster <- as_raster(rs)
analysis_series <- as_time_series(rs)
```

If an independently prepared field polygon is available, pass an `sf` or `sfc`
geometry to `mask`. This uses the supplied geometry directly and bypasses
delineation.

```{r supplied-mask, eval=FALSE}
# field_boundary is a user-supplied sf or sfc object
vi_raster <- as_raster(rs, mask = field_boundary)
vi_series <- as_time_series(rs, mask = field_boundary)
```

The mask may use any coordinate reference system, but it must have a defined
CRS and contain at least one non-empty geometry.

## Parameter settings

`rs_params()` returns the complete validated parameter list. Settings supplied
to `get_rs()` are stored in `rs$params` and provide the baseline for later
stages. A named argument supplied to a processing function overrides that
setting for the current call without changing the stored list.

```{r parameter-example}
example_params <- rs_params(
  perimeter_margins = c(0.30, 0.70),
  local_alive_density = 0.30
)

example_params[c("perimeter_margins", "local_alive_density")]
```

These are illustrative settings, not general recommendations. Parameter
effects and the four delineation stages are covered in the next vignette.

## Where to continue

- `vignette("ocular-build-pipeline")` explains field boundary delineation,
  direct stage calls, reusable pipelines, and multi-site processing.
- `vignette("ocular-experimental-features")` covers Landsat-MODIS data fusion,
  leave-one-out validation, and optional field boundary priors.
- `?get_rs`, `?rs_params`, and `?boundary_delineation` provide the complete API
  reference.
