---
title: "Experimental features"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Experimental features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(ocular)
```

`ocular` includes two experimental paths: Landsat-MODIS data fusion and an
optional field boundary prior. Both can affect an analysis and therefore need
validation for the locations, dates, inputs, and intended use. Neither feature
establishes general accuracy on its own.

The examples in this vignette require remote data or a user-supplied field
boundary file and are not evaluated when the package is built.

## Landsat-MODIS data fusion

`add_modis()` uses daily MCD43A4 NBAR data to estimate additional values for
vegetative indices (VI) on dates in the requested period that are not covered
by Landsat. A candidate MCD43A4 date is used only when no Landsat acquisition
falls within three days. The nearest Landsat acquisition provides the anchor,
and each estimate is created on the Landsat grid.

The appended scenes are estimates, not satellite observations. They extend
the output series but are excluded from the multi-temporal spectral feature
stack used for field boundary delineation.

### Adding estimated dates

Use an explicit Landsat source and a supported index. The location, dates, and
extent below are illustrative.

```{r add-fused-scenes, eval=FALSE}
rs <- get_rs(
  longitude = 147.227906,
  latitude = -33.0675167,
  start_date = "2016-05-18",
  end_date = "2016-09-21",
  index_name = "NDVI",
  source = "landsat-8",
  x_metres = 1000
)

rs_fused <- add_modis(rs)
field_fused <- boundary_delineation(rs_fused)
fused_series <- as_time_series(field_fused)
```

The time series includes `is_fused` so observed Landsat dates can be separated
from estimates. `anchor_mae` and `anchor_nmae` describe mismatch at the Landsat
anchor used to construct a fused scene. They are diagnostics for that
construction step, not held-out prediction errors.

`add_modis()` returns a Sentinel-2 object unchanged. It also leaves an object
unchanged when fusion has already been applied, no suitable MCD43A4 scenes are
available, or the selected index lacks the required Landsat and MCD43A4 band
mappings.

### Leave-one-out validation

`validate_data_fusion()` holds out each available Landsat scene in turn,
estimates it from the remaining inputs, and compares the estimate with the
observation. The standalone validation call must use the same output index as
the workflow being assessed; otherwise its default is EVI2.

```{r validate-fusion, eval=FALSE}
fusion_validation <- validate_data_fusion(
  147.227906,
  latitude = -33.0675167,
  start_date = "2016-05-18",
  end_date = "2016-09-21",
  index_name = "NDVI",
  source = "landsat-8",
  x_metres = 1000
)

fusion_validation
```

The result reports RMSE, MAE, bias, Pearson correlation, and supporting counts.
These values describe leave-one-out agreement for the selected query. They do
not demonstrate performance at other sites or periods and do not replace
comparison with independent data.

Validation returns `NULL` when fewer than two Landsat scenes or no usable
MCD43A4 scenes are available. Metric values are `NA` where fewer than
`min_valid_n` paired values support them; its default is four. Supporting
counts remain available for checking data sufficiency.

Fusion remains experimental. In particular, product-specific MCD43A4 quality
layers are not currently applied. Retain `is_fused` in downstream data, report
the validation conditions, and avoid presenting estimated dates as observed
Landsat acquisitions.

## Optional field boundary prior

`add_ftw_prior()` reads a user-supplied field boundary GeoParquet and selects
the polygon containing the supplied point. Fields of The World (FTW) is one
field boundary dataset that can be supplied in this form. fiboa is a schema,
not another name for the FTW dataset; the default geometry and date column
names follow that schema and can be overridden for another compatible export.

The selected polygon can supply calibration support. When it also has a usable
source year, it can apply an age-weighted soft prior during the initial area
search. An undated polygon provides calibration support without that
age-weighted prior.

### Attaching a prior

Retrieve the `ocular` object first, then provide the field boundary file and
its geometry CRS.

```{r attach-field-prior, eval=FALSE}
rs <- get_rs(
  longitude = 147.227906,
  latitude = -33.0675167,
  start_date = "2016-05-18",
  end_date = "2016-09-21",
  index_name = "NDVI",
  source = "landsat-8",
  x_metres = 1000
)

rs_prior <- add_ftw_prior(
  rs,
  source = "/path/to/field-boundaries.parquet",
  crs = 4326
)

field_prior <- boundary_delineation(rs_prior)
```

The default columns are `geometry` and `determination_datetime`. Use
`geometry_col` and `datetime_col` when the file uses different names; set
`datetime_col = NULL` when it has no observation date.

Local queries require the optional `duckdb` and `DBI` packages and a DuckDB
spatial extension that is already installed. Remote files additionally
require the local `httpfs` extension, and extracting a source year may require
`icu`. `ocular` does not install or download DuckDB extensions at runtime.

### Interpreting agreement diagnostics

`diagnose_against_ftw()` calculates pixel-level IoU, F1, precision, recall,
and supporting counts between a delineation and a reference polygon.

```{r diagnose-prior, eval=FALSE}
in_sample_diagnostic <- diagnose_against_ftw(field_prior)

# held_out_field_boundary is a user-supplied sf, sfc, or SpatVector object
independent_diagnostic <- diagnose_against_ftw(
  field_prior,
  reference = held_out_field_boundary
)
```

The first call compares the result with the same polygon that supplied the
prior. It is an in-sample diagnostic because that polygon may have influenced
the delineation. Use a separate held-out reference, as in the second call,
when assessing accuracy.

Whether an FTW or other compatible prior improves field boundary delineation
requires independent validation. Record the source file, schema mapping,
boundary date, and whether the comparison reference was also used as a prior.

## Before using an experimental result

- Validate the selected index, site, period, and input configuration.
- Keep observed and estimated dates distinguishable in fusion outputs.
- Treat agreement with a guiding field polygon as in-sample evidence.
- Report missing quality layers, missing boundary dates, and other limitations
  that affect interpretation.

See `?add_modis`, `?validate_data_fusion`, `?add_ftw_prior`, and
`?diagnose_against_ftw` for the complete interfaces.
