---
title: "Field boundary delineation and reusable pipelines"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Field boundary delineation and reusable pipelines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
library(ocular)
```

`boundary_delineation()` runs the default field boundary delineation workflow.
The four component functions can also be called directly when their order,
frequency, or settings need to be inspected or changed. A chosen sequence can
then be captured with `build_pipeline()` and applied to one or more retrieved
`ocular` objects.

Examples that retrieve satellite data are not evaluated during the vignette
build. Their locations, dates, and low-level parameter values are illustrative
rather than recommended settings.

## The four delineation stages

| Function | Role |
|---|---|
| `segment_area()` | Grows an initial candidate area from the supplied point using a constrained breadth-first search flood fill over a multi-temporal spectral feature stack. |
| `trace_perimeter()` | Refines an existing candidate boundary by removing or reviving perimeter cells according to local perimeter occupancy. When called before `segment_area()`, it stages one cleanup pass for the initial candidate. |
| `segment_interior()` | Reclassifies interior cells whose values for vegetative indices (VI) fall outside user-specified bounds or, during a divergent profile, differ from the seed signature. It does not alter perimeter cells. |
| `split_area()` | Looks for consecutive measurable line-of-sight profiles that diverge from the seed signature and rejects a provisional axis cut if it would remove any cell in the protected point-connected matching component. |

`boundary_delineation()` does more than call each stage once. It runs a fixed
sequence with repeated splitting and perimeter refinement;
`cleanup_boundary` controls the number of additional split-and-cleanup pairs.
Use `boundary_delineation()` unless the analysis requires a deliberately
different sequence.

## Running stages directly

First retrieve the scenes. The output index and the index used for delineation
can be different; this example requests NDVI output and otherwise retains the
default delineation index.

```{r retrieve-for-stages, eval=FALSE}
rs <- get_rs(
  longitude = 147.227906,
  latitude = -33.0675167,
  start_date = "2016-05-18",
  end_date = "2016-09-21",
  index_name = "NDVI",
  source = "landsat-8",
  x_metres = 1000
)
```

A direct chain makes each stage and its call-level settings explicit:

```{r direct-stage-chain, eval=FALSE}
candidate <- rs |>
  segment_area(
    area_sensitivity = 0.2,
    area_threshold = c(0.4, 1),
    local_alive_density = 0.25,
    search_windows = c(2L, 3L)
  )

candidate_raster <- as_raster(candidate)

field_manual <- candidate |>
  segment_interior(interior_sensitivity = 0.1) |>
  split_area(split_sensitivity = 0.05, split_gate = 2L) |>
  trace_perimeter(
    perimeter_margins = c(0.3, 0.9),
    trace_iter = 20L
  )
```

`candidate_raster` is the requested VI composite masked to the current
candidate area; it is not a binary-mask export. Comparing output after each
stage can help identify which part of a custom sequence needs adjustment.

`local_alive_density` is an admission condition in the initial flood fill: it
sets the required proportion of predicate-passing visited neighbours. It is
not a general thin-feature trimming control. `interior_sensitivity = NULL` or
`split_sensitivity = NULL` explicitly disables the corresponding stage.

## Stored and call-level parameters

Parameters supplied to `get_rs()` are stored on the returned object. A
possibly partial `params` list supplied to a later stage overlays those stored
settings for that call. Named stage arguments take precedence and do not
mutate `rs$params`.

```{r pipeline-settings}
pipeline_params <- rs_params(
  area_sensitivity = 0.2,
  area_threshold = c(0.4, 1),
  local_alive_density = 0.25,
  search_windows = c(2L, 3L),
  perimeter_margins = c(0.5, 0.9),
  trace_iter = 20L
)

pipeline_params[c(
  "area_sensitivity",
  "search_windows",
  "perimeter_margins"
)]
```

Calibration may provide per-window signatures and matching tolerances for the
initial area search. Without an attached field boundary prior, calibration uses
a complete seed-centred square and expands it until its precision condition is
met or the available feature grid limits further expansion. Explicit user
settings take precedence. `strictness` selects the robustness multiple applied
to measured per-window dispersion; its current engineering defaults require
external scientific validation.

## Preserving multiple areas

With `multiple_areas = FALSE`, output is limited to the centre-most connected
area (rank 1). Set it to `TRUE` when disconnected areas produced by the
seed-based search need to be retained separately.

```{r multiple-areas, eval=FALSE}
multi_params <- rs_params(multiple_areas = TRUE)

field_multi <- get_rs(
  longitude = -101.70862,
  latitude = 35.90839,
  start_date = "2010-04-25",
  end_date = "2010-10-15",
  x_metres = 1000,
  params = multi_params
) |>
  boundary_delineation()

multi_raster <- as_raster(field_multi)
multi_series <- as_time_series(field_multi)
```

When more than one disconnected area remains, raster layers are named
`area_1`, `area_2`, and so on, followed by `union`; the time series uses the
same per-area columns plus `union`. With one retained area, the ordinary
single-area output schema is used. These labels describe connected output
areas; they do not establish that each area is an independently validated
field.

## Capturing a reusable pipeline

`build_pipeline()` captures a supported stage chain that is applied to an
existing `ocular` object. It does not capture the initial `get_rs()` call:
retrieve an object first, then supply it when the pipeline is called. Omit that
initial object from the captured native pipe expression.

```{r construct-pipeline}
raster_pipeline <- build_pipeline(
  segment_area() |>
    trace_perimeter() |>
    as_raster(),
  params = pipeline_params
)

time_series_pipeline <- build_pipeline(
  boundary_delineation() |>
    as_time_series(),
  params = pipeline_params
)

class(raster_pipeline)
```

The pipeline can then be applied to a retrieved object:

```{r run-pipeline, eval=FALSE}
vi_raster <- raster_pipeline(rs)
```

Arguments that can be evaluated when the pipeline is constructed are stored
with it. A `params` list supplied when the pipeline is called replaces the
captured list for that run; stage arguments written directly in the captured
expression continue to take precedence.

```{r run-pipeline-override, eval=FALSE}
vi_raster <- raster_pipeline(
  rs,
  params = list(perimeter_margins = c(0.3, 0.7))
)
```

Supported captured stages are the four delineation functions,
`boundary_delineation()`, `add_modis()`, `as_raster()`, and
`as_time_series()`. Retrieval, `add_ftw_prior()`, and
`validate_data_fusion()` are separate operations.

## Applying a workflow across sites

`batch_rs()` accepts one row for each site and date range. Required columns are
`id`, `longitude`, `latitude`, `start_date`, and `end_date`.

```{r batch-input}
sites <- data.frame(
  id = c("field-a", "field-b"),
  longitude = c(147.2279, 147.3150),
  latitude = c(-33.0675, -33.1010),
  start_date = "2016-05-18",
  end_date = "2016-09-21"
)

sites
```

Without a custom pipeline, each row is retrieved, delineated with the default
workflow, and returned as a long time series.

```{r batch-default, eval=FALSE}
series <- batch_rs(
  sites,
  index_name = "NDVI",
  source = "landsat-8",
  x_metres = 1000,
  max_cloud_cover = 50
)
```

The result includes `id`, `index`, `source`, `date`, and `value`. A site that
fails is absent from the rows and recorded in `attr(series, "failures")`.
Retrieval is retried. After each uncached successful retrieval, the processing
pipeline is run once.

The batch interface currently exposes the scene-level `max_cloud_cover`
filter, but not the per-pixel `scl_classes` setting. Where per-pixel masking is
required, retrieve each site with `get_rs()` or prepare a separate controlled
workflow rather than assuming that the batch call applies it.

A custom batch pipeline must finish with a data frame containing `date` and
`value`. Caching is disabled for a custom pipeline unless a
`pipeline_cache_key` is supplied. Change that key whenever the pipeline or its
captured values change; reusing a key can reuse an earlier cached result.

```{r batch-custom, eval=FALSE}
series <- batch_rs(
  sites,
  pipeline = time_series_pipeline,
  pipeline_cache_key = "delineation-series-v1",
  index_name = "NDVI",
  source = "landsat-8",
  x_metres = 1000
)
```

## Where to continue

- `vignette("ocular-experimental-features")` covers Landsat-MODIS data fusion
  and compatible field boundary priors.
- `?rs_params` lists all implemented controls.
- `?build_pipeline` and `?batch_rs` document the reusable and multi-site
  interfaces.
